---
title: "Step H"
author: "Kora Welch"
date: "11/20/24"
output:
  html_document: default
  pdf_document: default
---
Download the dataset:
```{r}
library(readr)
ESS11 <- read_csv("ESS11.csv")
```

Hypotheses:
1) Education level influences one’s endorsement of women needing protection from men
2) Religiosity level influences one’s endorsement of women needing protection from men
3) With age, women will less likely endorse that women need protection from men, and men will more likely endorse that women need protection from men


Filter columns:
```{r}
library(tidyverse)

ess11New <- ESS11 %>%
  select(idno, gndr, wprtbym, rlgdgr, agea, edulvlb)
```

Re-name columns:
```{r}
colnames(ess11New) = c('Participant_id', 'Gender', 'Women_Protection', 'Religion_Level', 'Age', 'Highest_Edu_Lvl')
```

Remove/rename variables:
```{r}
ess11New <- subset(ess11New, ess11New$Age != 999)

ess11New <- subset(ess11New, ess11New$Highest_Edu_Lvl != 7777)
ess11New <- subset(ess11New, ess11New$Highest_Edu_Lvl != 8888)
ess11New <- subset(ess11New, ess11New$Highest_Edu_Lvl != 9999)

ess11New <- subset(ess11New, ess11New$Women_Protection != 7)
ess11New <- subset(ess11New, ess11New$Women_Protection != 8)
ess11New <- subset(ess11New, ess11New$Women_Protection != 9)

ess11New$Gender <- car::recode(ess11New$Gender, recodes = " '1' = '0' ; '2' = '1' ")
#0 = Female, 1 = Male
```


Data visualization:
```{r}
#Hypothesis 1
library(ggplot2)
library(jtools)

ggplot(ess11New, aes(Women_Protection, fill = as.factor(Highest_Edu_Lvl))) + geom_bar() + scale_color_manual() + theme_apa() + labs(x = "Belief that Women Need Protection From Men", y = "Frequency", title = "Education Level and Benevolent Sexism") + theme(text = element_text(family = "Times New Roman")) + scale_x_continuous(limits = c(0, 6), breaks = 1:5) + theme(strip.text.x = element_text(size = 12)) + theme(axis.text.x = element_text(size = 12)) + theme(axis.text.y = element_text(size = 12)) + theme(axis.title.x = element_text(size = 12)) + theme(axis.title.y = element_text(size = 12)) 
```


```{r}
#Hypothesis 2
library(ggplot2)
library(RColorBrewer)

ggplot(ess11New, aes(Religion_Level, fill = as.factor(Women_Protection))) + geom_bar() + labs(x = "Religion Level", y = "Count") + scale_fill_brewer(palette = "Reds") + theme_apa() + labs(x = "Religion Level", y ="Frequency", title = "Religion Level and Benevolent Sexism Endorsement") + theme(text = element_text(family = "Times New Roman")) +  theme(strip.text.x = element_text(size = 12)) + theme(axis.text.x = element_text(size = 12)) + theme(axis.text.y = element_text(size = 12)) + theme(axis.title.x = element_text(size = 12)) + theme(axis.title.y = element_text(size = 12)) + scale_x_continuous(limits = c(-1, 11), breaks = 0:10)
```


```{r}
#Hypothesis 3
library(ggplot2)
library(viridis)

ess11New$Gender <- as.factor(ess11New$Gender)
levels(ess11New$Gender)<- c("Female", "Male")

ggplot(ess11New, aes(Women_Protection, fill = as.factor(Age))) + geom_bar() + facet_wrap(~Gender) + scale_color_manual() + labs(x = "Belief that Women Need Protection From Men", y = "Frequency", title = "Genders' Benevolent Sexism Ratings and Age") + theme_apa() + theme(text = element_text(family = "Times New Roman")) + scale_x_continuous(limits = c(0, 6), breaks = 1:5) + theme(strip.text.x = element_text(size = 12)) + theme(axis.text.x = element_text(size = 12)) + theme(axis.text.y = element_text(size = 12)) + theme(axis.title.x = element_text(size = 12)) + theme(axis.title.y = element_text(size = 12))
```


Statistical analyses: 
```{r}
#Hypothesis 1: Education level influences one’s endorsement of women needing protection from men

ess11New$wp_num <- as.numeric(levels(as.factor(ess11New$Women_Protection)))[as.factor(ess11New$Women_Protection)]
is.numeric(ess11New$Women_Protection)

eduAnova <- aov(Women_Protection ~ as.factor(Highest_Edu_Lvl),  data = ess11New)
summary(eduAnova)

TukeyHSD(eduAnova)

etaSquared<- 1016/(1016+33833)
etaSquared
```


```{r}
#Hypothesis 2: Religiosity level influences one’s endorsement of women needing protection from men

wp_num <- as.numeric(levels(as.factor(ess11New$Women_Protection)))[as.factor(ess11New$Women_Protection)]
is.numeric(ess11New$Women_Protection)

religionAnova <- aov(as.numeric(Women_Protection) ~ as.factor(Religion_Level), data = ess11New)
summary(religionAnova)

TukeyHSD(religionAnova)

etaSquared2 <- 665/(665+34184)
etaSquared2
```


```{r}
#Hypothesis 3: With age, women will less strongly endorse that women need protection from men, and men will more strongly endorse that women need protection from men

ess11New$Gender <- as.factor(ess11New$Gender)
levels(ess11New$Gender) <- c(0, 1) 
# 0 = Female, 1 = Male
levels(ess11New$Gender)

ess11New$Gender <- relevel(ess11New$Gender, ref= "0") 

ageGenderlm <- lm(Women_Protection ~ Age * Gender, data = ess11New)
summary(ageGenderlm)
```



